## Chapter 6: Investigating the Waffles and Divorces
**6H1**. 
Use the Waffle House data, data(WaffleDivorce), to find the total causal influence of number of Waffle Houses on divorce rate. Justify your model or models with a causal graph.

```{r}
library(dagitty)
library(rethinking)

library(dagitty)

# New DAG with full variable names and new layout
waffle_dag <- dagitty('dag {
  Divorce [outcome, pos="2,1"]
  MedianAgeMarriage [pos="0,2"]
  MarriageRate [pos="1,1"]
  South [pos="-1,1.5"]
  WaffleHouses [exposure, pos="-1,0"]

  MedianAgeMarriage -> Divorce
  MedianAgeMarriage -> MarriageRate
  MarriageRate -> Divorce
  South -> MedianAgeMarriage
  South -> MarriageRate
  South -> WaffleHouses
  WaffleHouses -> Divorce
}')

drawdag(waffle_dag)

```
#### First we need to identify and block all backdoor paths. There are currently three backdoors from WaffleHouses to Divorce, that introduces the spurious associations:
WaffleHouses ← South → MedianAgeMarriage → Divorce
WaffleHouses ← South → Marriage → Divorce
Wafflehouses ← South → MedianAgeMarriageM → Marriage → Divorce
Since South is a cause of both WH and the downstream variables, conditioning on S, closes this backdoor and isolates the effect of Wafflehouses on Divoce rate. 



```{r}
data(WaffleDivorce)

WaffleDivorce$WaffleHouses_standardized <- scale(WaffleDivorce$WaffleHouses)
WaffleDivorce$Divorce_standardized <- scale(WaffleDivorce$Divorce)

m6H1 <- quap(
  alist(
    Divorce_standardized ~ dnorm(mu, sigma),
    mu <- a + bW * WaffleHouses_standardized + bS * South,
    a ~ dnorm(0, 0.2),
    bW ~ dnorm(0, 0.5),
    bS ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = WaffleDivorce
)

precis(m6H1)
```

### Causal effect of wafflehouses on divorce rate is very low with a mean around 0.1 and 89% interval lying in the range of -0.12 and 0.37. This indicates no causal effect between Wafflehouses and Divorce rate after blocking backdoors. 


**6H2**. 
Build a series of models to test the implied conditional independencies of the causal graph you used in the previous problem. If any of the tests fail, how do you think the graph needs to be amended? Does the graph need more or fewer arrows? Feel free to nominate variables that aren't int he data.

```{r}
impliedConditionalIndependencies(waffle_dag)
```
```{r}
data <- WaffleDivorce
data$AgeStd <- scale(data$MedianAgeMarriage)
data$MarriageStd <- scale(data$Marriage)
data$DivorceStd <- scale(data$Divorce)
data$WaffleStd <- scale(data$WaffleHouses)
data$SouthStd <- scale(data$South)


```


**Median Age at Marriage ⫫ Waffle Houses | South?**
```{r}
mod_age <- quap(
  alist(
    AgeStd ~ dnorm(mu, sigma),
    mu <- a + bWH * WaffleStd + bS * SouthStd,
    a ~ dnorm(0, 1),
    bWH ~ dnorm(0, 0.5),
    bS ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = data
)

plot(precis(mod_age))

```
### After conditioning on South, the association between WH and MAM becomes none. This confirms/supports the first of the DAG's implied conditional independencies


**Marriage Rate ⫫ Waffle Houses | South?**
```{r}
mod_mrate <- quap(
  alist(
    MarriageStd ~ dnorm(mu, sigma),
    mu <- a + bWH * WaffleStd + bS * SouthStd,
    a ~ dnorm(0, 0.5),
    bWH ~ dnorm(0, 0.5),
    bS ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = data
)

plot(precis(mod_mrate))
```
### Again the coefficient for WH is close to zero and supports with the assumption that Marriage Rate and Waffle Houses are independent after conditioning on South. 

**Divorce ⫫ South | Age, Marriage, Waffle Houses?**
```{r}
mod_divorce <- quap(
  alist(
    DivorceStd ~ dnorm(mu, sigma),
    mu <- a + bS * SouthStd + bAge * AgeStd + bM * MarriageStd + bWH * WaffleStd,
    a ~ dnorm(0, 0.5),
    bS ~ dnorm(0, 0.5),
    bAge ~ dnorm(0, 0.5),
    bM ~ dnorm(0, 0.5),
    bWH ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = data
)

plot(precis(mod_divorce))
```
### After conditioning on MedianAgeMarriageRate, MarriageRate and WaffleHouses, the coefficient for South becomes very small, and indicates a conditional independence between South and Divorce rate, when conditioning on the mediators.
